---
title: "Small Group Projects"
author: "Rylen Garlitz"
date: "`r Sys.Date()`"
output: html_document
---

# Install Libraries
```{r}
library(tidyverse)
#install.packages("readxlsb")
library(readxlsb)
```


## Read in Data

Data in XLSB is separated by year in each sheet. Using 'readxlsb' I read in each year as a separate file.

```{r}
#?read_xlsb
data_2023 <- read_xlsb("data/2007-2023-PIT-Counts-by-CoC.xlsb", sheet = 1)
data_2022 <- read_xlsb("data/2007-2023-PIT-Counts-by-CoC.xlsb", sheet = 2)
data_2021 <- read_xlsb("data/2007-2023-PIT-Counts-by-CoC.xlsb", sheet = 3)
data_2020 <- read_xlsb("data/2007-2023-PIT-Counts-by-CoC.xlsb", sheet = 4)
data_2019 <- read_xlsb("data/2007-2023-PIT-Counts-by-CoC.xlsb", sheet = 5)
data_2018 <- read_xlsb("data/2007-2023-PIT-Counts-by-CoC.xlsb", sheet = 6)
data_2017 <- read_xlsb("data/2007-2023-PIT-Counts-by-CoC.xlsb", sheet = 7)
data_2016 <- read_xlsb("data/2007-2023-PIT-Counts-by-CoC.xlsb", sheet = 8)
data_2015 <- read_xlsb("data/2007-2023-PIT-Counts-by-CoC.xlsb", sheet = 9)
data_2014 <- read_xlsb("data/2007-2023-PIT-Counts-by-CoC.xlsb", sheet = 10)
data_2013 <- read_xlsb("data/2007-2023-PIT-Counts-by-CoC.xlsb", sheet = 11)
```

### Read only Gainesville/Alachua, Putnam County Data

```{r}
data_2023 <- data_2023 %>% 
  filter(CoC.Name == "Gainesville/Alachua, Putnam Counties CoC")
data_2022 <- data_2022 %>% 
  filter(CoC.Name == "Gainesville/Alachua, Putnam Counties CoC")
data_2021 <- data_2021 %>% 
  filter(CoC.Name == "Gainesville/Alachua, Putnam Counties CoC")
data_2020 <- data_2020 %>% 
  filter(CoC.Name == "Gainesville/Alachua, Putnam Counties CoC")
data_2019 <- data_2019 %>% 
  filter(CoC.Name == "Gainesville/Alachua, Putnam Counties CoC")
data_2018 <- data_2018 %>% 
  filter(CoC.Name == "Gainesville/Alachua, Putnam Counties CoC")
data_2017 <- data_2017 %>% 
  filter(CoC.Name == "Gainesville/Alachua, Putnam Counties CoC")
data_2016 <- data_2016 %>% 
  filter(CoC.Name == "Gainesville/Alachua, Putnam Counties CoC")
data_2015 <- data_2015 %>% 
  filter(CoC.Name == "Gainesville/Alachua, Putnam Counties CoC")
data_2014 <- data_2014 %>% 
  filter(CoC.Name == "Gainesville/Alachua, Putnam Counties CoC")
data_2013 <- data_2013 %>% 
  filter(CoC.Name == "Gainesville/Alachua, Putnam Counties CoC")
```

### Add year column

```{r}
data_2023$Year <- 2023
data_2022$Year <- 2022
data_2021$Year <- 2021
data_2020$Year <- 2020
data_2019$Year <- 2019
data_2018$Year <- 2018
data_2017$Year <- 2017
data_2016$Year <- 2016
data_2015$Year <- 2015
data_2014$Year <- 2014
data_2013$Year <- 2013
```

### Merge Data sets

```{r}
#?merge
data <- merge(data_2023, data_2022, all = TRUE)
data <- merge(data, data_2021, all = TRUE)
data <- merge(data, data_2020, all = TRUE)
data <- merge(data, data_2019, all = TRUE)
data <- merge(data, data_2018, all = TRUE)
data <- merge(data, data_2017, all = TRUE)
data <- merge(data, data_2016, all = TRUE)
data <- merge(data, data_2015, all = TRUE)
data <- merge(data, data_2014, all = TRUE)
data <- merge(data, data_2013, all = TRUE)
```

### Graph some stuff

```{r}
#Overall Homeless Count
ggplot(data = data, mapping = aes(x = Year,y = Overall.Homeless)) + 
geom_line()

#Homeless 18-24
ggplot(data = data, aes(x = Year, y = Overall.Homeless...Age.18.to.24)) +
  geom_line()

#Homeless Under 18
ggplot(data = data, aes(x = Year, y = Overall.Homeless...Under.18)) + 
  geom_line() + 
  labs(y = "Total Homeless Under 18")

```




### Total Sheltered vs Unsheltered
```{r sheltered unsheltered}
#Total Sheltered Homeless
ggplot(data = data, aes(x = Year, y = Sheltered.Total.Homeless)) + 
  geom_line() + 
  labs(y = "Total Sheltered Homeless")

#Total Unsheltered Homeless
ggplot(data = data, aes(x = Year, y = Unsheltered.Homeless)) + 
  geom_line() + 
  labs(y = "Total Unsheltered Homeless")

#Total Unsheltered vs Total Sheltered.
data %>% 
ggplot(aes(x = Year)) + 
  geom_line(aes(y = Sheltered.Total.Homeless, color = "Sheltered")) + 
  geom_line(aes(y = Unsheltered.Homeless, color = "Unsheltered")) +
  labs(y = "Number of Homeless")


#ES - Emergency Shelter
#TH - Transitional Housing
#SH - Safe Haven
```

What happened in 2015? A HUGE decrease in the total number of unsheltered homeless. Nearly the same number as sheltered. 




### Male vs. Female Homeless numbers

```{r male v. fem}
#Male vs. Female
ggplot(data = data, aes(x = Year)) + 
  geom_line(aes(y = Overall.Homeless...Female, color = "Female")) + 
  geom_line(aes(y = Overall.Homeless...Male, color = "Male")) + 
  labs(y = "Number of Homeless")

```

2015 again?? Male homelessness decreased rapidly and female homelessness had a slight increase! 

Trend of males having higher rates of homelessness that females.

### Overall Homeless by Race

```{r race plot}
ggplot(data = data, aes(x = Year)) + 
  geom_line(aes(y = Overall.Homeless...White, color = "White")) + 
  geom_line(aes(y = Overall.Homeless...Black..African.American..or.African, color = "Black, African American, African")) +
  geom_line(aes(y = Overall.Homeless...Asian.or.Asian.American, color = "Asian, Asian American")) + 
  geom_line(aes(y = Overall.Homeless...American.Indian..Alaska.Native..or.Indigenous, color = "American Indian, Alaskan, Indigenous")) +
  labs(y = "Overall Homeless")
```

Whites are the majority of the homeless followed by Blacks (2015 AGAIN?). Asian and Inigenous homelessness is very low. 



